Hybrid platforms and bargaining power

### 2023-12-06 src/paper/application.tex
Update proofs and propositions for general case

Finally, let us conclude this section by establishing a stronger result for a more restrictive class of profit functions.
In the following, I assume that the profit function has a specific, additive form.
Intuitively, this means that -- in terms of total profits generated -- the platform's and the fringe firms' products are substitutable to each other according to some constant ratio.\footnote{
    Note that it does not mean that their products are substitutes in the usual, demand theory sense.
    They still might be complements to each other.
}
\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let the weight function be constant ($w(t) \equiv 1$).
    % TODO: I think the weight function assumption is not necessary.
    Furthermore, let $f$ have the following, additive form:
    \begin{align*}
        f(N_P, N_F) = g(\alpha N_P + \beta N_F)
    \end{align*}
    where $\alpha, \beta > 0$.
    Also assume that $g$ is twice differentiable and that $g'' < 0$.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
\end{proposition}
As opposed to the previous results, this proposition tells us something about the total profits generated by the platform and the fringe firms as a function of the platform's product variety, and under equilibrium entry behavior by the fringe firms.
In particular, it states that total profits increase if the total profit function is convex (products are substitutes), and decreases otherwise.
This result has particularly strong implications in models where total profits (or total product variety) has a monotone relationship with consumer welfare, such as those in \textcite{anderson2020aggregative}.
\Cref{sec:example} also presents a model in which this is the case.

@@

\begin{proof}[Proof of \cref{prop:share_of_fringe}]
    Remember that
    \begin{align*}
        \pi_F(N_P, N_F) = f(N_P, N_F) - \int_0^1 w(s) f(N_P, N_F) \ds.
    \end{align*}
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} &= \frac{\partial}{\partial N_P} \left[ f(N_P, N_F) - \int_0^1 w(s) f(N_P, N_F) \ds \right]\\
        &= \frac{\partial f(N_P, N_F)}{\partial N_P} - \int_0^1 w(s) \frac{\partial f(N_P, sN_F)}{\partial N_P} \ds \\
        &= \underbrace{h(N_F)}_{(*)} - \underbrace{\int_0^1 w(s) h(s N_F)}_{(**)}
    \end{align*}
    where $h(N_F) = \frac{\partial h(N_F)}{\partial N_P}$.

    Notice that $(*)$ and $(**)$ are both positive.
    Furthermore, $(**)$ is the weighed average of $(*)$ over the interval $[0, N_F]$.
    Therefore, if $h(N_F)$ is strictly increasing, then $(*) > (**)$, and the whole expression is positive.
    Conversely, it is negative if $h(N_F)$ is strictly decreasing.

    If $h(N_F)$ is differentiable, then it is equivalent to saying that
    \begin{align*}
        \frac{\mathrm{d}h(N_F)}{\mathrm{d}N_F} = \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} < 0 &\implies \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} < 0 \\
        \frac{\mathrm{d}h(N_F)}{\mathrm{d}N_F} = \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} > 0 &\implies \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} > 0
    \end{align*}
\end{proof}

\begin{proof}[Proof of \cref{cor:fringe_entry}]
    \Cref{prop:unique_equilibrium} establishes that if an equilibrium with $N_F > 0$ exists, it is unique.
    Use the implicit function theorem on the equation from \cref{ass:free_entry} to obtain
    \begin{align*}
        \frac{\partial N_F}{\partial N_P} = \frac{\frac{\partial \pi_F(N_P, N_F)}{\partial N_P}}{I_F - \frac{\partial \pi_F (N_P, N_F)}{\partial N_F}}
    \end{align*}
    The derivative exists if the above expression is well-defined, i.e. $\frac{\partial \pi_F (N_P, N_F)}{\partial N_F} \neq I_F$.
    Remember from \cref{prop:share_of_fringe} that the numerator of this expression is negative under the condition $\frac{\partial^2 f(N_P, N_F)}{\partial N_P \partial N_F}$.
    Now all we need to show to conclude the proof is that
    \begin{align*}
        \frac{\partial \pi_F (N_P, N_F)}{\partial N_F} < I_F,
    \end{align*}
    which is the statement of \cref{lem:slope_at_eq}.
\end{proof}

@@

    First, let us show that fringe profits are lower than total investment costs if $\bar N_F = N_F^*(N_P') - \frac{\alpha}{\beta}\Delta$.
    In order to do this, consider this formulation of fringe profits:
    \begin{align*}
        \pi_F(N_P', N_F') &= f(N_P', N_F') - \int_0^1 w(s) f(N_P', t N_F') \dt \\
        &= \int_0^1 t \frac{\partial f(N_P', n_F')}{\partial n_F} \bigg|_{n_F = t N_F'},
    \end{align*}
    which can be obtained by integration by parts.
    Now let us use the functional form assumption $f(N_P, N_F) = g(\alpha N_P + \beta N_F)$ to obtain
    \begin{align*}
        \pi_F(N_P, N_F) = \int_0^1 t \underbrace{\frac{\partial}{\partial t}g(\alpha N_P + \beta t N_F)}_{\coloneq D(N_P, N_F)} \dt.
    \end{align*}

    Let us focus on the term $D(N_P, N_F)$.
    First, calculate
    \begin{align*}
        D(N_P', N_F') = \beta N_F' g'(\alpha N_P' + \beta t N_F').
    \end{align*}
    Let us do the same for $D(N_P'', \bar{N}_F) = D(N_P' + \Delta, N_F - \frac{\alpha}{\beta} \Delta)$ to get
    \begin{align*}
        D(N_P'', \bar{N}_F) = \beta \bar{N}_F g'(\alpha N_P' + \beta t N_F' + \underbrace{\Delta \alpha (1-t)}_{> 0}).
    \end{align*}
    Substitute these expressions into $\pi_F$ to get
    \begin{align*}
        \pi_F(N_P', N_F') &= N_F' \beta \int_0^1 t g'(\alpha N_P' + \beta t N_F') \dt, \\
        \pi_F(N_P'', \bar{N}_F) &= \bar{N}_F \beta \int_0^1 t g'(\alpha N_P' + \beta t N_F' + \Delta \alpha (1-t)) \dt.
    \end{align*}
    Finally, let us bound the difference between $\pi_F(N''_Pm \bar{N}_F) - I_F \bar{N}_F$ from above:
    \begin{align*}
        \pi_F(N_P'', \bar{N}_F) - I_F \bar{N}_F &= \frac{\bar{N}_F}{N_F'}\left[\beta \int_0^1 t g'(\alpha N_P' + \beta t N_F' + \Delta \alpha (1-t)) \dt - I_F N_F' \right] \\
        &< \frac{\bar{N}_F}{N_F'}\left[\beta \int_0^1 t g'(\alpha N_P' + \beta t N_F') \dt - I_F N_F' \right] \\
        &= \frac{\bar{N}_F}{N_F'}\left[ \pi_F(N_P', N_F') - I_F N_F \right] \\
        &= 0.
    \end{align*}
    The inequality is due to the concavity of $g$, and the final equality is the result of profits being zero in the equilibrium $(N_P', N_F')$.

    All that remains to be shown is that $\pi_F(N_P'', N_F'') < N_F'' I_F$.
    This is a consequence of the fact that, under \cref{ass:single_crossing}, $\pi_F$ is either concave or decreasing in $N_F$.
    As a consequence, if it at some point ($\bar{N}_F$) dips below the linear function $N_F \mapsto I_F N_F$, it can never reach it again for any $N_F'' > \bar{N}_F$.
    With this we arrive at a contradiction, as $(N_P'', N_F'')$ cannot be an equilibrium.

>>>>

Finally, let us conclude this section by establishing a stronger result for a more restrictive class of profit functions.
In the following, I assume that the profit function has a specific, additive form.
Intuitively, this means that -- in terms of total profits generated -- the platform's and the fringe firms' products are substitutable to each other according to some constant ratio.\footnote{
    Note that it does not mean that their products are substitutes in the usual, demand theory sense.
    They still might be complements to each other.
}
\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let $f$ have the following, additive form:
    \begin{align*}
        f(N_P, N_F) = g(\alpha N_P + \beta N_F)
    \end{align*}
    where $\alpha, \beta > 0$.
    Also assume that $g$ is twice differentiable and that $g'' < 0$.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
\end{proposition}
As opposed to the previous results, this proposition tells us something about the total profits generated by the platform and the fringe firms as a function of the platform's product variety, and under equilibrium entry behavior by the fringe firms.
In particular, it states that total profits increase if the total profit function is convex (products are substitutes), and decreases otherwise.
This result has particularly strong implications in models where total profits (or total product variety) has a monotone relationship with consumer welfare, such as those in \textcite{anderson2020aggregative}.
\Cref{sec:example} also presents a model in which this is the case.

@@

\begin{proof}[Proof of \cref{prop:share_of_fringe}]
    Remember that
    \begin{align*}
        \pi_F(N_P, N_F, N_C) = N_C \int_0^1 w(s) N_F \partial_2 f(N_P, s N_F) \ds,
    \end{align*}
    where $w(s)$ is either $s$ or $s^2$, depending on whether one assumes one or two-sided bargaining.
    By assumption, $f$ is twice continuously differentiable, therefore $\partial_2$ is also continuously differentiable in $N_P$.
    Thus the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} &= \frac{\partial}{\partial N_P} N_C \int_0^1 w(s) N_F \partial_2 f(N_P, s N_F) \ds \\
        &= N_C \int_0^1 w(s) N_F \frac{\partial}{\partial N_P} \partial_2 f(N_P, s N_F) \ds \\
        &= N_C \int_0^1 w(s) N_F \partial^2_{12} f(N_P, s N_F) \ds.
    \end{align*}

    As $w(s) > 0$ for $s > 0$, if $\partial_2 f(N_P, s N_F)$ has the same sign over $[0, N_F]$, then the integral also has the same sign.
    Formally,
    \begin{align*}
        \forall n_F \in [0, N_F] \enspace \frac{\partial^2 f(N_P, n_F)}{\partial N_P \partial n_F} < 0 &\implies \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} < 0 \\
        \forall n_F \in [0, N_F] \enspace \frac{\partial^2 f(N_P, n_F)}{\partial N_P \partial n_F} > 0 &\implies \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} > 0
    \end{align*}
\end{proof}

\begin{proof}[Proof of \cref{cor:fringe_entry}]
    \Cref{prop:unique_equilibrium} establishes that if an equilibrium with $N_F > 0$ exists, it is unique.
    Use the implicit function theorem on the equation from \cref{ass:free_entry} to obtain
    \begin{align*}
        \frac{\partial N_F}{\partial N_P} = \frac{\frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P}}{I_F - \frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F}}
    \end{align*}
    The derivative exists if the above expression is well-defined, i.e. $\frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F} \neq I_F$.
    Remember from \cref{prop:share_of_fringe} that the numerator of this expression is negative under the condition $\frac{\partial^2 f(N_P, N_F, N_C)}{\partial N_P \partial N_F}$.
    Now all we need to show to conclude the proof is that
    \begin{align*}
        \frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F} < I_F,
    \end{align*}
    which is the statement of \cref{lem:slope_at_eq}.
\end{proof}

@@

    First, let us show that fringe profits are lower than total investment costs if $\bar N_F = N_F^*(N_P') - \frac{\alpha}{\beta}\Delta$.
    In order to do this, consider the following formulation of the fringe profit function:
    \begin{align*}
        \pi_F(N_P', N_F', N_C) &= \int_0^1 w(t) \frac{\Pi(N_P, t N_F, N_C)}{\partial t} \dt \\
        &= N_C \int_0^1 w(t) \underbrace{\frac{\partial}{\partial t}g(\alpha N_P + \beta t N_F)}_{\coloneq D(N_P, N_F)} \dt,
    \end{align*}
    where $w(t) = t$ or $w(t) = t^2$ in the one-sided and two-sided bargaining cases, respectively.

    Let us focus on the term $D(N_P, N_F)$.
    First, calculate
    \begin{align*}
        D(N_P', N_F') = \beta N_F' g'(\alpha N_P' + \beta t N_F').
    \end{align*}
    Let us do the same for $D(N_P'', \bar{N}_F) = D(N_P' + \Delta, N_F - \frac{\alpha}{\beta} \Delta)$ to get
    \begin{align*}
        D(N_P'', \bar{N}_F) = \beta \bar{N}_F g'(\alpha N_P' + \beta t N_F' + \underbrace{\Delta \alpha (1-t)}_{> 0}).
    \end{align*}
    Substitute these expressions into $\pi_F$ to get
    \begin{align*}
        \pi_F(N_P', N_F', N_C) &= N_C N_F' \beta \int_0^1 w(t) g'(\alpha N_P' + \beta t N_F') \dt, \\
        \pi_F(N_P'', \bar{N}_F, N_C) &= N_C \bar{N}_F \beta \int_0^1 w(t) g'(\alpha N_P' + \beta t N_F' + \Delta \alpha (1-t)) \dt.
    \end{align*}
    Finally, let us bound the difference between $\pi_F(N''_Pm \bar{N}_F) - I_F \bar{N}_F$ from above:
    \begin{align*}
        \pi_F(N_P'', \bar{N}_F, N_C) - I_F \bar{N}_F &= \frac{\bar{N}_F}{N_F'}\left[\beta N_C \int_0^1 w(t) g'(\alpha N_P' + \beta t N_F' + \Delta \alpha (1-t)) \dt - I_F N_F' \right] \\
        &< \frac{\bar{N}_F}{N_F'}\left[\beta N_C \int_0^1 w(t) g'(\alpha N_P' + \beta t N_F') \dt - I_F N_F' \right] \\
        &= \frac{\bar{N}_F}{N_F'}\left[ \pi_F(N_P', N_F', N_C) - I_F N_F \right] \\
        &= 0.
    \end{align*}
    The inequality is due to the concavity of $g$, and the final equality is the result of profits being zero in the equilibrium $(N_P', N_F', N_C)$.

    All that remains to be shown is that $\pi_F(N_P'', N_F'', N_C) < N_F'' I_F$.
    This is a consequence of the fact that, under \cref{ass:single_crossing}, $\pi_F$ is either concave or decreasing in $N_F$.
    Therefore, if it at some point ($\bar{N}_F$) dips below the linear function $N_F \mapsto I_F N_F$, it can never reach it again for any $N_F'' > \bar{N}_F$.
    With this we arrive at a contradiction, as $(N_P'', N_F'', N_C)$ cannot be an equilibrium.
